Checking Invariant Confluence, In Whole or In Parts \thanks{ The original version of this paper is entitled ``Interactive Checks for Coordination Avoidance'' and was published in PVLDB Vol.\ 12. }

### 2020-05-22 doc/sigmod_paper/sections/abstract.tex
Polished up paper, now 8 pages.
Cleaned up commented out text.

In this paper, we establish conditions under which a commonly used sufficient
condition for \invariantconfluence{} is both necessary and sufficient, and we
use this condition to design a general-purpose interactive
\invariantconfluence{} decision procedure. We then take a step beyond
\invariantconfluence{} and introduce a generalization of
\invariantconfluence{}, called segmented \invariantconfluence{}, that allows us
to replicate non-\invariantconfluent{} objects with a small amount of
coordination.
%
We implemented these formalisms in a prototype called Lucy and found that our
decision procedures efficiently handle common real-world workloads including
%
% foreign keys, rollups, escrow transactions, and more.
foreign keys, escrow transactions, and more.
%
We also found that segmented \invariantconfluent{} replication can deliver up
to an order of magnitude more throughput than linearizable replication.
% for low contention workloads and comparable throughput for medium to high
% contention workloads.
\end{abstract}

>>>>

In this paper, we establish conditions under which a commonly used sufficient
condition for \invariantconfluence{} is both necessary and sufficient, and we
use this condition to design a general-purpose interactive
\invariantconfluence{} decision procedure. We then take a step beyond
\invariantconfluence{} and introduce a generalization of
\invariantconfluence{}, called segmented \invariantconfluence{}, that allows us
to replicate non-\invariantconfluent{} objects with a small amount of
coordination. We implemented these formalisms in a prototype called Lucy and
found that our decision procedures efficiently handle common real-world
workloads including foreign keys, escrow transactions, and more.
\end{abstract}

### 2020-05-22 doc/sigmod_paper/sections/conclusion.tex
Cleaned up commented out text.

\section{Conclusion}
This paper revolved around two major contributions.
%
First, we developed a deeper understanding of \invariantclosure{} and
\invariantconfluence{} by looking at the two criteria with reachability in
mind. We found that \invariantclosure{} fails to incorporate a notion of
reachability, and using this intuition, we developed conditions under which
\invariantclosure{} and \invariantconfluence{} are equivalent. We implemented
this insight in an interactive \invariantconfluence{} decision procedure
that automatically checks whether an object is \invariantconfluent{}, with the
assistance of a programmer.

Second, we proposed a new consistency model and generalization of
\invariantconfluence{}, segmented \invariantconfluence{}, that can be used to
replicate non-\invariantconfluent{} objects with a small amount of coordination
while still preserving their invariants. We found that segmented
\invariantconfluence{} naturally subsumes existing techniques for maintaining
invariants of replicated objects (e.g.\ locking and escrow transactions), and
we developed an interactive decision procedure for segmented
\invariantconfluence{}.

Through our evaluation, we found that our decision procedures could analyze a
number of realistic workloads, each in less than a second. We also showed that
segmented \invariantconfluence{} can significantly outperform linearizable
replication for low-coordination workloads.

>>>>

\section{Conclusion}
This paper revolved around two major contributions. First, we found that
\invariantclosure{} fails to incorporate a notion of reachability, and using
this intuition, we developed conditions under which \invariantclosure{} and
\invariantconfluence{} are equivalent. We implemented this insight in an
interactive \invariantconfluence{} decision procedure that automatically checks
whether an object is \invariantconfluent{}, with the assistance of a
programmer.
%
Second, we proposed a new consistency model and generalization of
\invariantconfluence{}, segmented \invariantconfluence{}, that can be used to
replicate non-\invariantconfluent{} objects with a small amount of coordination
while still preserving their invariants.

### 2020-05-22 doc/sigmod_paper/sections/invariant_confluence.tex
Joe feedback.
Cleaned up commented out text.

% Note that by modelling a transaction $t$ as a function $S \to S$, we focus
% exclusively on the effects that a transaction has on the object (i.e.\
% ``writes'' to the object). Transactions are also free to read the value of the
% object, but these reads are not captured by our model because, as we'll see,
% they do not affect \invariantconfluence{}. For example, we could model any
% read-only transaction as a function $t$ where $t(s) = s$ for every $s \in S$.

@@

Second, a server $p_i$ can send its state $s_i$ to another server $p_j$ with
state $s_j$ causing $p_j$ to transition from state $s_j$  to state $s_i \join
s_j$. Servers periodically merge states with one another in order to keep their
states loosely synchronized.
% \footnote{%
%   Notably, if $O$ is a CRDT---i.e. $O$ is a semilattice and every transaction
%   $t \in T$ is inflationary---then this periodic merging ensures that $O$ is
%   strongly eventually consistent~\cite{shapiro2011conflict}.
% }.
Note that unlike with transactions, servers \emph{cannot} abort a merge; server
$p_j$ must transition from $s_j$ to $s_i \join s_j$ whether or not $s_i \join
s_j$ satisfies the invariant.

@@

We say an expression $e$ is \defword{\sTIreachable{}} if it corresponds to a
valid execution of our system model. Formally, we define
$\sTIreachablepredicate{e}$ to be the smallest predicate that satisfies the
following equations:
\begin{itemize}
  \item
    $\sTIreachablepredicate{s_0}$.
  \item
    For all expressions $e$ and for all transactions $t$ in the set $T$ of
    transactions, if $\sTIreachablepredicate{e}$ and $I(t(e))$, then
    $\sTIreachablepredicate{t(e)}$.
  \item
    For all expressions $e_1$ and $e_2$, if $\sTIreachablepredicate{e_1}$ and
    $\sTIreachablepredicate{e_2}$, then $\sTIreachablepredicate{e_1 \join
    e_2}$.
\end{itemize}
Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists an
\sTIreachable{} expression $e$ that evaluates to $s$. Returning to
\exampleref{Ints} with start state $s_0 = 42$, we see that all integers greater
than or equal to 42 (i.e.\ $\setst{x \in \ints}{x \geq 42}$) are
\sTIreachable{}, and all other integers are \sTIunreachable{}.

>>>>

@@

Second, a server $p_i$ can send its state $s_i$ to another server $p_j$ with
state $s_j$ causing $p_j$ to transition from state $s_j$ to state $s_i \join
s_j$. Servers periodically merge states with one another in order to keep their
states loosely synchronized. Note that unlike with transactions, servers
\emph{cannot} abort a merge; server $p_j$ must transition from $s_j$ to $s_i
\join s_j$ whether or not $s_i \join s_j$ satisfies the invariant.

@@

We say an expression $e$ is \defword{\sTIreachable{}} if it corresponds to a
valid execution of our system model. Formally, we define
$\sTIreachablepredicate{e}$ to be the predicate that satisfies the
following conditions:
\begin{itemize}
  \item
    $\sTIreachablepredicate{s_0}$.
  \item
    For all expressions $e$ and for all transactions $t$ in the set $T$ of
    transactions, if $\sTIreachablepredicate{e}$ and $I(t(e))$, then
    $\sTIreachablepredicate{t(e)}$.
  \item
    For all expressions $e_1$ and $e_2$, if $\sTIreachablepredicate{e_1}$ and
    $\sTIreachablepredicate{e_2}$, then $\sTIreachablepredicate{e_1 \join
    e_2}$.
\end{itemize}
Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists an
\sTIreachable{} expression $e$ that evaluates to $s$. Returning to
\exampleref{Ints} with start state $s_0 = 42$, we see that all integers greater
than or equal to 42 (i.e.\ $\setst{x \in \ints}{x \geq 42}$) are
\sTIreachable{}, and all other integers are \sTIunreachable{}.
